---
title: "Visualizing Data"
author: "Ismail Fadeli"
date: "August 19, 2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Visualization in R
In order to be able to visualize data, we have to install \textbf{tidyverse} package in R Studio. Here's the code for that:
```{r tidyverse}
library(tidyverse)
```
## Viewing Dataset
To visualize cars miles per gallon dataset from the USA datacenter:
```{r mpg}
mpg
```

To plot \textbf{mpg} data, we need to run this code to put \textbf{displ} into x-axis and \textbf{hwy} into the y-axis.

```{r pressure}
  ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy))
```
The visualized graph shows a negative relationship between the engine size and fuel efficiency. Bigger engine size uses more fuel than lower engine sizes to travel the same distance.

With \textbf{ggplot}, you begin a plot with the function \textbf{ggplot()} creates a coordinate system that you can add layers to. The first argument of \textbf{ggplot} is the dataset to use in the graph. So \textbf{ggplot(data = mpg)} creates an empty graph.
We can complete the graph by adding more layers to  \textbf{ggplot()}. The function \textbf{geompoint()} creates a layer to your plot, which creates a scatterplot.
The mapping argument is always paired with \textbf{aes()}, and the x and y arguments of aes() specify which variables to map to the x-axes and y-axes.

## Exercice
1. When running the code \textbf{ggplot(data = mpg)}, we see an empty graph:
2. The \textbf{mtcars} dataset has 32 rows and 11 columns.
```{r}
mtcars

```
3. The \textbf{drv} variable describes the type of the car, meaning it is either a front wheel drive, rear wheel, or four wheel drive.

4. The code for a scatter plot of \textbf{hwy} versus \textbf{cyl}:
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cyl, y = hwy))
```

## Aesthetic Mappings
You can show information about your data by mapping the aesthetics in your plot to the variables in your dataset. For example, you can map the colors of your points to the class variable to reveal the class of each car:
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

In the previous example, we mapped the class to the color aesthetic, but we could have mapped class to the size aesthetic in the same way. In this case, the exact size of each point would reveal its class affiliation. We get a warning here, because mapping an unordered variable (class) to an ordered variable (size) is not a good idea:
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```
Or we could have mapped class to the \textbf{alpha} aesthetic, which controls transparency of the points, or the shape of the points:
```{r}
# Top
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))

# Buttom
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```
You can also set the aesthetic properties of your geom manually. For example, we can make all of the points in our plot blue:
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```
## Facets
one way to split categorical variables is to plot into facets, subplots that each display one subset of the data.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~class, nrow = 2)
```
To facet the plot on combination of two variables, add \textbf{facetgrid} to your plot call. The first argument of \textbf{facetgrid} is also a formula. This time the formula should contain two variable names separated by a ~:
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ cyl)
```
## Geometric Objects
```{r}
# Left
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
# Right
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))

```
```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )
```
```{r}
ggplot(mpg) +
  geom_text(mapping = aes(x = displ, y = hwy, label = model))
```

In order to display multiple geoms in the same plot, add multiple geom functions to \textbf{ggplot()}:
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy)) 

```
```{r}
ggplot(data = mpg, mapping = aes( x = displ, y = hwy)) +
  geom_point() +
  geom_smooth() 

  
```
```{r}
ggplot(data = mpg, mapping = aes( x = displ, y = hwy)) +
  geom_point(mapping = aes(color = class)) +
  geom_smooth() 

```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = class)) +
  geom_smooth(
    data = filter(mpg, class == "subcompact"),
    se = FALSE
  )
  
```
```{r}
diamonds
```
## Statistical Transformation
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```
```{r}
ggplot(data = diamonds) +
  stat_count(mapping = aes(x = cut))
```
You might want to override the default mapping from trans-
formed variables to aesthetics. For example, you might want to
display a bar chart of proportion, rather than count:
```{r}
ggplot(data = diamonds) +
  geom_bar(
    mapping = aes(x = cut, y = ..prop.., group = 1)
  )

```
You might want to draw greater attention to the statistical trans-
formation in your code. For example, you might use stat_sum
mary(), which summarizes the y values for each unique x value,
to draw attention to the summary that you're computing
```{r}
ggplot(data = diamonds) +
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```
##Position Adjustments
You can color a bar chart using either the color aesthetic, or more usefully, fill:
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, color = cut))
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))
```
Note what happens if you map the fill aesthetic to another vari-
able, like clarity: the bars are automatically stacked. Each colored
rectangle represents a combination of cut and clarity:
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity))
```

position = "identity" will place each object exactly where it
falls in the context of the graph. This is not very useful for bars,
because it overlaps them. To see that overlapping we either need
to make the bars slightly transparent by setting alpha to a small
value, or completely transparent by setting fill = NA:
```{r}
ggplot(
  data = diamonds,
  mapping = aes(x = cut, fill = clarity)
) +
  geom_bar(alpha = 1/5, position = "identity")
ggplot(
  data = diamonds,
  mapping = aes(x = cut, color = clarity)
)
```
position = "fill" works like stacking, but makes each set of
stacked bars the same height. This makes it easier to compare
proportions across groups:
```{r}
  ggplot(data = diamonds) +
  geom_bar(
    mapping = aes(x = cut, fill = clarity),
    position = "fill"
  )
```
position = "dodge" places overlapping objects directly beside
one another. This makes it easier to compare individual values:
```{r}
ggplot(data = diamonds) +
  geom_bar(
    mapping = aes(x = cut, fill = clarity), 
    position = "dodge"
  )
```
here's one other type of adjustment that's not useful for bar charts,
but it can be very useful for scatterplots. Recall our first scatterplot.
Did you notice that the plot displays only 126 points, even though
there are 234 observations in the dataset?

The values of hwy and displ are rounded so the points appear on a
grid and many points overlap each other. This problem is known as
overplotting. This arrangement makes it hard to see where the mass
of the data is. Are the data points spread equally throughout the
graph, or is there one special combination of hwy and displ that
contains 109 values

You can avoid this gridding by setting the position adjustment to
"jitter." position = "jitter" adds a small amount of random noise
to each point. This spreads the points out because no two points are
likely to receive the same amount of random noise:
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy),
    position = "jitter"
    )
```

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
  geom_boxplot()
```
```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
  geom_boxplot() +
  coord_flip()
```
```{r}
nz <- map_data("nz")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", color = "black") 
  
```
coord_quickmap() sets the aspect ratio correctly for maps. This
is very important if you're plotting spatial data with ggplot2
(which unfortunately we don't have the space to cover in this
book):
```{r}
ggplot(nz, aes(long, lat, group = group))+
  geom_polygon(fill = "white", color = "black") +
  coord_quickmap()
```
coord_polar() uses polar coordinates. Polar coordinates reveal
an interesting connection between a bar chart and a Coxcomb
chart:
```{r}
bar <- ggplot(data = diamonds) +
  geom_bar(
    mapping = aes(x = cut, fill = cut),
    show.legend = FALSE,
    width = 1
  ) +
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)
  
  bar + coord_flip()
  bar + coord_polar()
```
```{r}
my_val <- 10
my_val
```

