---
title: "Data Transformation with dplyr"
author: "Ismail Fadeli"
date: "August 27, 2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Transformation with dplyr

```{r}
library(nycflights13)
```

```{r }
library(tidyverse)
```
```{r}
flights
```
If we want to see all the data in flights: View(flights)

## Filter Rows with filter()
\textbf{filter()} allows you to subset observations based on their values. The first argument is the name of the dataframe. The second and the rest arguments are the expressions that filter the dataframe.
For example, we can select all flights on January 1st with:

```{r}
filter(flights, month == 1, day ==1)
```

To assign the filtered result into a new dataframe we can use an assignement:
```{r}
jan1 <- filter(flights, month == 1, day == 1)
```
If you want to print the results and save them in a variable at the same time, you can use parentheses:
```{r}
(dec25 <- filter(flights, month ==12, day == 25))
```
## Comparisons
The standard operations that R uses are: >, >=, <, <=, != (not equal), and == (equal).
```{r}
sqrt(2) ^ 2 == 2
```
```{r}
1/49 *49 == 1
```
```{r}
near(sqrt(2) ^ 2, 2)
```

```{r}
near(1 / 49 * 49, 1)
```

Boolean operators in R are: & is "and"", | is "or"", ! is "not". The following code shows the months of November or December:
```{r}
filter(flights, month == 11 | month == 12)
```

```{r}
nov_dec <- filter(flights, month %in% c(11, 12))
```
```{r}
filter(flights, !(arr_delay > 120 | dep_delay > 120))
filter(flights, arr_delay <= 120, dep_delay <= 120)
```

## Missing Values

if you want to determine if a value is missing, use it.na():

```{r}
x <- NA
```
```{r}
is.na(x)
```
```{r}
df <- tibble(x = c(1, NA, 3))
filter(df, x > 1)
```

```{r}
filter(df, is.na(x) | x > 1)
```
Finding all flights that flew to Houston:
```{r}
flights
```

```{r}
filter(flights, dest == "IAH" | dest == "HOU")
```

```{r}
View(flights)
```
Finding all flights that were operated by United, American, or Delta:
```{r}
  filter(flights, carrier == "UA" | carrier == "AA" | carrier == "DL")
```
```{r}
filter(flights, between(x = month, left = 7, right = 9))
```
##Arrange Rows with Arrange()
Arrange() takes the dataframe and a set of column names to order by.
```{r}
arrange(flights, year, month, year)
```
Use desc() to reorder by a column in descending order:
```{r}
arrange(flights, desc(arr_delay))
```
Missing values are always sorted at the end:
```{r}
df <- tibble(x = c(5, 2, NA))
arrange(df, x)
```
```{r}
arrange(df, desc(x))
```

## Select Columns with select()
```{r}
# Select columns by name
select(flights, year, month, day)
```

```{r}
# Select all columns between year and day
select(flights, year:day)
```
```{r}
# Select all columns except those from year to day
select(flights, -(year:day))
```
There are a number of helper functions you can use within select():
starts_with("abc") matches names that begin with "abc".
. ends_with("xyz") matches names that end with "xyz".
. contains("ijk") matches names that contain "ijk".
. matches("(.)\\1") selects variables that match a regular
expression. This one matches any variables that contain
repeated characters. You'll learn more about regular expressions
in Chapter 11.
. num_range("x", 1:3) matches x1, x2, and x3.


Use rename() to rename a variables:
```{r}
rename(flights, tail_num = tailnum)
```

Another option is to use select() in conjuction with the everything() helper. This is usefull if we had a number of variables we wanted to move to the start of the dataframe.
```{r}
select(flights, time_hour, air_time, everything())
```

## Exercice: 
1. Selecting dep_time, dep_delay, arr_time, arr_delay from flights:
```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
```
```{r}
select(flights, dep_time, dep_delay, dep_delay)
```

```{r}
select(flights, contains("TIME"))
```

## Adding new variables with mutate()
mutate() always adds new columns to the end of the dataset, so let's start by making a small dataset and adding a column using mutat():
```{r}
flights_small <- select(flights, 
  year:day,
  ends_with("delay"),
  distance,
  air_time              
)
```

```{r}
View(flights_small)
```
```{r}
mutate(flights_small,
  gain = arr_delay - dep_delay,
  speed = distance / air_time * 60
  )
```

```{r}
Sales <- read_csv("D:/USPOLOASSN/Sales/SALESSS.csv")
```
```{r}
Sales
```
```{r}
ggplot(data = Sales) + 
  geom_bar(mapping = aes(x = R))
  
```
```{r}
sales_par_category <- read_csv("D:/USPOLOASSN/Sales/Sales_by_cat.csv")
```

```{r}
ggplot(data = sales_par_category) +
  geom_bar(mapping = aes(x = Category, fill = Category)) +
    coord_flip() 

```
```{r}
select(sales_par_category, "Category")
```
```{r}
mutate(sales_par_category, "NEW")
```
```{r}
sales_par_category
```
```{r}
sales_par_category %>%
  count(Category)
```

```{r}
sales_par_category
```
```{r}
sales_par_category %>%
  group_by(Category) %>%
  count()
```
```{r}
problems(sales_par_category)
```
```{r}
library(data.table)
```

```{r}
fread()
```
```{r}
S <- (sales_par_category %>%
  group_by(Category) %>%
  count())
```
```{r}
S
```
```{r}
max(S$n)
```
```{r}
pp <-arrange(S, desc(n))

```
```{r}
pp
```
```{r}
ma <- filter(pp, n == max(pp$n))
ma
```

```{r}
i = pp[1][1]
 sss=      pp[2]
```
```{r}
mutate(i)
```


